---
title: 'Homework 3'
subtitle: "PSTAT 120C: Design and Analysis of Experiments" 

author: "Instructor: Ali Abuzaid"
format: 
 pdf:
    code-fold: true
    code-line-numbers: true
    code-copy: true
    code-tools: true
    self-contained: true
    toc: false
    toc-location: left
    number-sections: true
    geometry: margin=0.7in 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message =  FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(error =  FALSE)
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


:::callout
<div style="text-align: center">
<span style="color: blue;"> **STUDENT NAME **</span>  
</div>
- Brandon Wong II (3621406)


:::

:::{.callout-tip}
# Instructions

* This homework aims to familiarize you with real-life data.

* Please write your **full name** and **perm number** legibly on the front page of your homework sheet.

* This homework must be completed and submitted **individually**. Collaboration is allowed for discussion, but each student must submit their own work.

* Solve all homework problems. You may use `R` for your calculations and analysis, but you must include all `R` code used.  

* Your reasoning must be clear and complete. To receive full credit, provide sufficient explanations alongside any code and results.

* Ensure that all `R` code, mathematical formulas, and workings are presented clearly and appropriately.

* All figures should be numbered, and axes must be labeled. 

* Please use the provided `HW 3 S26.qmd` file to type your solutions and submit the completed assignment as a PDF file. You can utilize `RStudio` for this purpose. For guidance, refer to the [Tutorial: Hello, Quarto](https://quarto.org/docs/get-started/hello/rstudio.html)).

* Submit your solution via **Gradescope** as **a single PDF file**.


:::

:::{.callout-caution}
# Due Date
**Due Date:** Friday, May 8, 2026 at 11:59 PM
:::


:::{.callout-caution}
# Note to Students
Solve all the questions both manually and using `R` commands, and ensure that your final answers from both methods are clearly presented and compared.
:::



\newpage

`r bfcolor("ANSWER ALL THE FOLLOWING QUESTIONS:", "blue")` \

:::{.callout-important}
## Question 1  
 Do you hate Mondays? Researchers in Germany have provided another reason for you: They
concluded that the risk of heart attack on a Monday for a working person may be as much as
50% greater than on any other day.1 The researchers kept track of heart attacks and coronary
arrests over a period of 5 years among 330,000 people who lived near Augsberg, Germany. In an
attempt to verify the researcher’s claim, 200 working people who had recently had heart attacks
were surveyed. The day on which their heart attacks occurred appear in the following table.

\centering
\begin{tabular}{ccccccc}
\hline
Sunday & Monday & Tuesday & Wednesday & Thursday & Friday & Saturday \\
24 & 36 & 27 & 26 & 32 & 26 & 29 \\
\hline
\end{tabular}
Do these data present sufficient evidence to indicate that there is a difference in the percentages
of heart attacks that occur on different days of the week? Test using $\alpha = 0.05$.

- Obtain the p-value.

:::


`r bfcolor("ANSWERS TO QUESTION 1:", "red")` \

$$H_0: p_{Sun}=p_{Mon}=p_{Tue}=p_{Wed}=p_{Thu}=p_{Fri}=p_{Sat}= 1/7$$
$$H_a: \text{at least one } p_i \ne 1/7$$
```{r}
n <- 200
p <- 1/7

exp <- n*p
obs <- c(24, 36, 27, 26, 32, 26, 29)

diff <- obs - exp
diff

chisq <- sum(diff^2)/exp
chisq

df <- length(obs) - 1

# Critical Value
crit_val <- qchisq(0.05, df, lower.tail = FALSE)
crit_val

crit_val > chisq # ==> Don't reject H0 if TRUE, Reject H0 if FALSE

# Finding p-value
p_val <- pchisq(chisq, df, lower.tail = FALSE)
p_val
```

**At a significance level $\alpha$ = 0.05, we do not have enough evidence to conclude that there is a difference in the percentages of heart attacks that occur on different days of the week.**







\newpage
:::{.callout-important}
## Question 2  
The data in the following table are the frequency counts for 400 observations on the number
of bacterial colonies within the field of a microscope, using samples of milk film.2 Is there
sufficient evidence to claim that the data do not fit the Poisson distribution? (Use $\alpha = 0.05$.)
	
\textbf{Note:} For the `R` solution, you may use any appropriate built-in function where applicable, **and also** provide your own code to perform the calculations.

\centering
\begin{tabular}{cc}
\hline
\textbf{Number of Colonies per Field} & \textbf{Frequency of Observation} \\
\hline
0  & 56 \\
1  & 104 \\
2  & 80 \\
3  & 62 \\
4  & 42 \\
5  & 27 \\
6  & 9 \\
7  & 9 \\
8  & 5 \\
9  & 3 \\
10 & 2 \\
11 & 0 \\
19 & 1 \\
\hline
   & 400 \\
\hline
\end{tabular}

:::

`r bfcolor("ANSWERS TO QUESTION 2:", "red")` \


```{r}
# Inputting Data
counts <- c(rep(0, 56), rep(1, 104), rep(2, 80), rep(3, 62), 
            rep(4, 42), rep(5, 27), rep(6, 9), rep(7, 9), rep(8, 5), 
            rep(9, 3), rep(10, 2), rep(11, 0), rep(19, 1))

freq <- table(counts)

obs_freq <- c(freq["0"], freq["1"], freq["2"], freq["3"], freq["4"],
         freq["5"], freq["6"], freq["7"], freq["8"], freq["9"],
         freq["10"], 0, freq["19"])

result <- data.frame(
  Yi = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "19"),
  Frequency = as.numeric(obs_freq))

result

# Estimating Parameter (MLE of Poisson: lambda_hat = Y_bar)
n <- sum(obs_freq)
y <- as.numeric(result$Yi)

lambda_MLE <- sum(y * result$Frequency) / n # lambda <- sum(counts) / length(counts); yields the same
# lambda_MLE

# Estimating Expected Probabilities under H0
expected_prob <- dpois(y, lambda_MLE)
# expected_prob

# Expected Frequency (or counts)
options(scipen = 999)
expected_freq <- expected_prob * n
# expected_freq

# Difference between observed frequency an expected frequency
freq_diff <- obs_freq - expected_freq
# freq_diff

# Pearson Chi-Sq Test
chisq_2 <- sum((freq_diff^2) / expected_freq)
chisq_2


dpois(19, lambda_MLE)*400

```





\newpage
:::{.callout-important}
## Question 3 
 Does education really make a difference in how much money you will earn? Researchers randomly
selected 100 people from each of three income categories—“marginally rich,” “comfortably
rich,” and “super rich”—and recorded their education levels. The data is summarized
in the table that follows.


\centering
\begin{tabular}{lccc}
\hline
\textbf{Highest Education Level} & \textbf{Marginally Rich} & \textbf{Comfortably Rich} & \textbf{Super Rich} \\
\hline
No college            & 32 & 20 & 23 \\
Some college          & 13 & 16 & 1  \\
Undergraduate degree  & 43 & 51 & 60 \\
Postgraduate study    & 12 & 13 & 16 \\
\hline
Total & 100 & 100 & 100 \\
\hline
\end{tabular}


\begin{enumerate}
	\item Describe the independent multinomial populations whose proportions are compared in the
$\chi^2$ analysis.

\item Do the data indicate that the proportions in the various education levels differ for the three
income categories? Test at the $\alpha = 0.01$ level.

\end{enumerate}

:::
`r bfcolor("ANSWERS TO QUESTION 3:", "red")` \

**Q1:**

Education Level: No College, Some College, Undergraduate Degree, Postgraduate Study
Income Categories: Marginally Rich, Comfortably Rich, Super Rich

**Q2:**

$$H_0: \text{The proportion of the 3 education levels is the same across the 3 income categories (Homogenous)}$$
$$H_a: \text{The proportion in the 3 education levels differ across the 3 income categories (Not Homogenous)}$$

```{r}
data <- matrix(c(32, 20, 23,
                 13, 16, 1,
                 43, 51, 60,
                 12, 13, 16),
               nrow = 4, byrow = TRUE)

colnames(data) <- c("Marginally_Rich", "Comfortably_Rich", "Super_Rich")
rownames(data) <- c("No_College", "Some_College", "Undergraduate", "Postgraduate")

data
```





\newpage

:::{.callout-important}
## Question 4 

The results of a study suggest that the initial electrocardiogram (ECG) of a suspected heart
attack victim can be used to predict in-hospital complications of an acute nature. The study
included 469 patients with suspected myocardial infarction (heart attack). Each patient was
categorized according to whether their initial ECG was positive or negative and whether the
person suffered life-threatening complications subsequently in the hospital. The results are
summarized in the following table.


\centering
\begin{tabular}{cccc}
\hline
 & \multicolumn{2}{c}{\textbf{Subsequent In-Hospital Life-Threatening Complications}} & \\
\cline{2-3}
\textbf{ECG} & \textbf{No} & \textbf{Yes} & \textbf{Total} \\
\hline
Negative & 166 & 1  & 167 \\
Positive & 260 & 42 & 302 \\
\hline
Total & 426 & 43 & 469 \\
\hline
\end{tabular}


Is there sufficient evidence to indicate that whether or not a heart attack patient suffers
complications depends on the outcome of the initial ECG? Test using $\alpha = 0.05$.
:::

`r bfcolor("ANSWERS TO QUESTION 4:", "red")` \



```{r}

```











